Exploring Nonlinear Atmospheric Influences on Temperature Using Polynomial Regression at Vancouver International Airport

### 2024-11-30 paper/paper.qmd
update data and discussion section

The limitations are documented in @sec-methodology.

@@

### Potential Directions for Expanded Analysis
Expanding this research to include airports in different regions around the world would provide a broader understanding of how variables influence airport safety. Airports operate in different climatic and geographic conditions such as coastal locations and high-altitude terrains, have unique challenges influenced by local weather patterns. Investigating these variations could show patterns and trends that are not observable when focusing on Vancouver's airport.

>>>>

### Sources of Bias
The dataset used in this study, derived from Environment and Climate Change Canada (ECCC) and the Canadian Centre for Climate Services contains adjustments and homogenization to ensure consistency and quality. Adjustments for factors such as station relocations, changes in measurement instruments, and the missing data can introduce localized variations that do not reflect natural atmospheric conditions. The detailed sources of bias and limitations are documented in @sec-methodology.

@@

### Using the Model to Predict Temperature Effects on Airport Operations

The relationship between mean temperature and airport operations, particularly aircraft takeoff performance shows the importance of predictive modeling in adapting to the impacts of climate change. Rising temperatures affect air density, reducing lift and necessitating longer runways or weight restrictions for safe takeoff [@coffel_climate_2015]. These changes not only compromise operational efficiency but also increase costs for airports and airlines. As climate change accelerates, mean temperature fluctuations will likely become more frequent and extreme, posing greater challenges for airport operations.

One key application of the model lies in runway management and expansion. Airports facing persistent increases in mean temperature can use predictions from the model to assess the need for longer runways. These expansions would accommodate higher takeoff speeds required under low air density conditions. Similarly, the model's outputs can inform flight scheduling and weight management strategies by identifying periods of elevated temperatures when weight restrictions may be necessary. This allows airports to prioritize lighter-loaded flights, optimize cargo and passenger distribution, and adjust scheduling to minimize disruptions.

The model also supports integration into broader climate resilience strategies. Airports can combine temperature predictions with long-term climate adaptation plans, enabling investments in infrastructure and technology suited for a warming climate. Moreover, by embedding the model into real-time decision support systems, airports can dynamically adjust operations based on updated atmospheric data.

### Potential Directions for Expanded Analysis

Expanding this research to include airports in different regions around the world would provide a broader understanding of how variables influence airport safety. Airports operate in different climatic and geographic conditions such as coastal locations and high-altitude terrains, have unique challenges influenced by local weather patterns. Investigating these variations could show patterns and trends that are not observable when focusing on Vancouver's airport.
